Chemical Tagging Using Minimum Message Length. I.\\ Gaussian Mixture Modelling

### 2017-04-24 articles/chemical-tagging-gmm/ms.tex
some bad text

\begin{equation}
I(\weights) = \frac{K - 1}{2}\log{N} - \frac{1}{2}\sum_{k=1}^{K}\weight_k - \log{\Gamma{\left(K\right)}} \quad ,
\end{equation}

@@

% The fisher information matrix

\subsection{Prior probabilities on the parameters}
\label{sec:priors}

\subsection{Expected Fisher information}

Computing the expected Fisher information requires the second order partial
derivatives of the negative log-likelihood $-\likelihood(\data|\vecmean,\veccov)$.
This is difficult to calculate for mixtures of Gaussians, and for this reason
we approximate the determinant of the Fisher information 
$|\mathcal{F}(\vecmean, \veccov)|$ as the product of 
$|\mathcal{F}\left(\vecmean\right)|$ and $|\mathcal{F}\left(\veccov\right)|$ 
\citep{Oliver_1996,Roberts_1998}. 
We begin by taking second derivative of the log-likelihood function
$\likelihood\left(\data|\vecmean,\veccov\right)$,

@@

\begin{eqnarray}
  |\mathcal{F}(\vecmean)| = \left|-\frac{\delta^2\likelihood}{\delta\vecmean^2}\right| = N^d|\veccov|^{-1} \quad .
\end{eqnarray}

@@

\subsection{Encoding the message}

Given our priors, likelihood function, and approximate expressions for the
Fisher information, we are now in a position to calculate the total message
length.
There are a number of components that go into the message:

A common method to estimate the parameters of a Gaussian distribution is by
maximum likelihood, where the log-likelihood $\likelihood$ is maximized. In
order to do this, one computes the gradient of the log-likelihood function
with respect to the parameters and solves the resulting equations.

- Encoding the message.

- Process of expectation and maximization for each step.

@@

This research is supported by an Australian Research Council Discovery Project
grant (DP160100637). 
Source code for this project is available at \texttt{\giturl}. This document
was compiled on \gitdate\ from revision hash \texttt{\githash} in that
repository.

>>>>

\begin{equation}
I(\weights) = \frac{K - 1}{2}\log{N} - \frac{1}{2}\sum_{k=1}^{K}\log\weight_k - \log{\Gamma{\left(K\right)}} \quad ,
\end{equation}

@@

Computing the expected Fisher information $\mathcal{F}$ requires the second 
order partial derivatives of the negative log-likelihood 
$-\likelihood(\data|\vecmean,\veccov)$.
We approximate the determinant of the Fisher information 
$|\mathcal{F}(\vecmean_k, \veccov_k)|$ as the product of 
$|\mathcal{F}\left(\vecmean_k\right)|$ and $|\mathcal{F}\left(\veccov_k\right)|$ 
\citep{Oliver_1996,Roberts_1998}. 
We begin by taking second derivative of the log-likelihood function
$\likelihood\left(\data|\vecmean,\veccov\right)$ for a multivariate Gaussian
distribution,

@@

\begin{eqnarray}
  |\mathcal{F}(\vecmean_k)| = \left|-\frac{\delta^2\likelihood}{\delta\vecmean^2}\right| = N^d|\veccov|^{-1} \quad .
\end{eqnarray}

@@

Thus for all $K$ mixtures, Equation \ref{eq:10} expands to:

\begin{eqnarray}
I(\vecmean, \veccov) &=& -\sum_{k=1}^K\frac{D + 1}{2}\log{|\veccov_k|} 
                         + \frac{1}{2}\sum_{k=1}^K\left[\frac{D(D+3)}{2}\log{N} - D\log{2} - (D + 2)\log{|\veccov_k|}\right] \nonumber \\
I(\vecmean, \veccov) &=& -\sum_{k=1}^K\frac{D + 1}{2}\log{|\veccov_k|}
                         + \frac{KD(D + 3)}{4}
                         + \frac{D\log{2}}{2}
                         - \frac{1}{2}\sum_{k=1}^K (D+2)\log{|\veccov_k|} \nonumber\\
I(\vecmean, \veccov) &=& -\sum_{k=1}^K(D + 1)\log{|\veccov_k|} + \frac{1}{2}\left(\frac{KD(D + 3)}{2} + D\log{2}\right)
\end{eqnarray}

\subsubsection{Encoding the data, given the model parameters}
\label{sec:encoding-data}

Each datum can only be stated with a precision which is given by the accuracy
of the measurement.
If we assume homoskedastic noise properties for the observed data, and the
precision that each elemental abundance can be measured is $\epsilon$, then
the \emph{probability} of a datum is given by 
$\textrm{Pr}(\datum_n) = \epsilon^D\textrm{Pr}(\datum_i|\mathcal{M})$
where $\mathcal{M}$ represents the model and $\textrm{Pr}(\datum_i|\mathcal{M})$
is the \emph{probability density}.
Note that $\epsilon$ is a constant, and is only used to maintain validity 
between \emph{probability} and \emph{probability density}.
For these experiments we adopt $\epsilon = 0.01$, but note that the value of
$\epsilon$ has no effect on our inferences.

The length of the total encoding of a datum given the model is,
\begin{equation}
  I(\data_n) = -\log{\textrm{Pr}(\data_n)} = -D\log\epsilon - \log\sum_{k=1}^{K}w_{k}f_{k}(\data_n|\vecmean_k,\veccov_k)
\end{equation}

\noindent{}and for the entire data:

\begin{equation}
  I(\data|\vectheta) = -ND\log\epsilon - \sum_{n=1}^{N}\log\sum_{k=1}^{K}w_{k}f_k(\data_n|\vecmean_k,\veccov_k)
\end{equation}

\subsubsection{Encoding the lattice constant}

\subsection{The Objective Function}
\label{sec:objective-function}

The message lengths of individual components allows us to fully specify an
objective function to minimise the total message length, given any number of
Gaussian mixtures $K$ and parameter estimates $\vectheta$:

\begin{equation}
% TODO:
  I(\vectheta,\data) = I(K) + I(\weights) + \sum_{k=1}^{K}I(\vectheta_k) + I(\data|\vectheta) + \textrm{lattice} \quad .
\end{equation}

Substituting the expressions outlined in previous subsections, we arrive at:

\begin{equation}
  % FULL EQUATION.
  I(\vectheta,\data) = long
\end{equation}

Note that our objective function includes the number of mixtures $K$, a
discrete valued parameter. 
Because $K$ is not continuous, this complicates our optimization procedure.

Here we will outline our optimization procedure to minimise the message length
when $K$ is known and fixed, before briefly describing our search strategy to
move between trial values of $K$.

@@

This research has made use of NASA's Astrophysics Data System.
This research is supported by an Australian Research Council Discovery Project
grant (DP160100637). 
Source code for this project is available at \texttt{\giturl}. This document
was compiled on \gitdate\ from revision hash \texttt{\githash} in that
repository.
